---
title: "Main Analysis"
output: html_document
date: "2025-07-17"
---

# 0 Set up environment
```{r}
# Clean environment
rm( list = ls() )

# set path 
path <- "~/GitHub/GBS-GWAS-Methods-Sensitivity-Analyses/Empirical Analysis/inputs/"
outpath <- "~/GitHub/GBS-GWAS-Methods-Sensitivity-Analyses/Empirical Analysis/outputs/"
  

# load libraries
library(qqman)
library(circlize)
library(VennDiagram)
library(stringr)
library(tidyverse)
library(ggrepel)
library(ggplot2)

# Prepare Plotting functions 
QQ_plot <- function(data, type){
  df <- data.frame(expected= (-log10(ppoints(length(data |> pull(lrt.pvalue))))), 
                             observed=(-log10(sort(data |> pull(lrt.pvalue))))) 
  df |>
    ggplot(aes(x = expected, y = observed)) +
  geom_point() +
  geom_abline(slope = 1, color="red", lty="dashed")  +      
  labs(
    x = "-log10(Expected P)",
    y = "-log10(Observed P)",
    title = paste0(type," QQ plot")
  ) +
  theme_minimal()+
   ylim(0, max(c(df$expected, df$observed)))+
   xlim(0, max(c(df$expected, df$observed)))+ 
    ggtitle("")
}

# Load table
Missingness_Prop <- read.delim(paste0(path,"Missingness_Prop.txt"), header=F, sep=" ")
colnames(Missingness_Prop) <- c("POS", "Missingness_Proportion")
```

# 1. SENSITIVTY ANALYSIS : POPULATION STRUCTURE CORRECTION METHODS

# 1.1. CC
```{r}
## SNP-GWAS
# Load GWAS results
SNPGWAS_CC_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/CC/SNPGWAS.txt"), sep = "\t", header=T)|> 
  filter(af>=0.05, af <0.95)
# Significant hits
SNPGWAS_CC_sighits <- SNPGWAS_CC_data |>  filter(lrt.pvalue < 0.05 /nrow(SNPGWAS_CC_data)) 
# QQ plot 
SNPGWAS_CC_qqplot <- QQ_plot(SNPGWAS_CC_data, "SNP-GWAS") 

## DBGWAS
# Load GWAS results
DBGWAS_CC_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/CC/DBGWAS.txt"), sep = "\t", header=T) |>  
  filter(af >=0.05, af <0.95)
# Significant hits
DBGWAS_CC_sighits <- DBGWAS_CC_data |> filter(lrt.pvalue < 0.05 /87899)  
# QQ plot 
DBGWAS_CC_qqplot <- QQ_plot(DBGWAS_CC_data, "DBGWAS")

## PanGWAS
# Load GWAS results
PanGWAS_CC_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/CC/PanGWAS.txt"), sep = "\t", header=T) |> filter(af >=0.05, af <0.95)
# Get significant hits
PanGWAS_CC_sighits <- PanGWAS_CC_data |> filter(lrt.pvalue < 0.05 / nrow(PanGWAS_CC_data)) 
# QQ plot 
PanGWAS_CC_qqplot <- QQ_plot(PanGWAS_CC_data, "PanGWAS")
```
# 1.2. SC
```{r}
## SNP-GWAS
# Load GWAS results
SNPGWAS_SC_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/SC/SNPGWAS.txt"), sep = "\t", header=T)|> 
  filter(af>=0.05, af <0.95)
# Significant hits
SNPGWAS_SC_sighits <- SNPGWAS_SC_data |>  filter(lrt.pvalue < 0.05 /nrow(SNPGWAS_SC_data)) 
# QQ plot 
SNPGWAS_SC_qqplot <- QQ_plot(SNPGWAS_SC_data, "SNP-GWAS") 

## DBGWAS
# Load GWAS results
DBGWAS_SC_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/SC/DBGWAS.txt"), sep = "\t", header=T) |>  
  filter(af >=0.05, af <0.95)
# Significant hits
DBGWAS_SC_sighits <- DBGWAS_SC_data |> filter(lrt.pvalue < 0.05 /87899)  
# QQ plot 
DBGWAS_SC_qqplot <- QQ_plot(DBGWAS_SC_data, "DBGWAS")

## PanGWAS
# Load GWAS results
PanGWAS_SC_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/SC/PanGWAS.txt"), sep = "\t", header=T) |> filter(af >=0.05, af <0.95)
# Get significant hits
PanGWAS_SC_sighits <- PanGWAS_SC_data |> filter(lrt.pvalue < 0.05 / nrow(PanGWAS_SC_data)) 
# QQ plot 
PanGWAS_SC_qqplot <- QQ_plot(PanGWAS_SC_data, "PanGWAS")
```

# 1.3. mash_distances
```{r}
## SNP-GWAS
# Load GWAS results
SNPGWAS_mash_distances_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/Mash_distances/SNPGWAS.txt"), sep = "\t", header=T)|> 
  filter(af>=0.05, af <0.95)
# Significant hits
SNPGWAS_mash_distances_sighits <- SNPGWAS_mash_distances_data |>  filter(lrt.pvalue < 0.05 /nrow(SNPGWAS_mash_distances_data)) 
# QQ plot 
SNPGWAS_mash_distances_qqplot <- QQ_plot(SNPGWAS_mash_distances_data, "SNP-GWAS") 

## DBGWAS
# Load GWAS results
DBGWAS_mash_distances_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/Mash_distances/DBGWAS.txt"), sep = "\t", header=T) |>  
  filter(af >=0.05, af <0.95)
# Significant hits
DBGWAS_mash_distances_sighits <- DBGWAS_mash_distances_data |> filter(lrt.pvalue < 0.05 /87899)  
# QQ plot 
DBGWAS_mash_distances_qqplot <- QQ_plot(DBGWAS_mash_distances_data, "DBGWAS")

## PanGWAS
# Load GWAS results
PanGWAS_mash_distances_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/Mash_distances/PanGWAS.txt"), sep = "\t", header=T) |> filter(af >=0.05, af <0.95)
# Get significant hits
PanGWAS_mash_distances_sighits <- PanGWAS_mash_distances_data |> filter(lrt.pvalue < 0.05 / nrow(PanGWAS_mash_distances_data)) 
# QQ plot 
PanGWAS_mash_distances_qqplot <- QQ_plot(PanGWAS_mash_distances_data, "PanGWAS")
```

# 1.4. SNP_kinship
```{r}
## SNP-GWAS
# Load GWAS results
SNPGWAS_SNP_kinship_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/SNP_kinship/SNPGWAS.txt"), sep = "\t", header=T)|> 
  filter(af>=0.05, af <0.95)
# Significant hits
SNPGWAS_SNP_kinship_sighits <- SNPGWAS_SNP_kinship_data |>  filter(lrt.pvalue < 0.05 /nrow(SNPGWAS_SNP_kinship_data)) 
# QQ plot 
SNPGWAS_SNP_kinship_qqplot <- QQ_plot(SNPGWAS_SNP_kinship_data, "SNP-GWAS") 

## DBGWAS
# Load GWAS results
DBGWAS_SNP_kinship_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/SNP_kinship/DBGWAS.txt"), sep = "\t", header=T) |>
  filter(af >=0.05, af <0.95)
# Significant hits
DBGWAS_SNP_kinship_sighits <- DBGWAS_SNP_kinship_data |> filter(lrt.pvalue < 0.05 /87899)  
# QQ plot 
DBGWAS_SNP_kinship_qqplot <- QQ_plot(DBGWAS_SNP_kinship_data, "DBGWAS")

## PanGWAS
# Load GWAS results
PanGWAS_SNP_kinship_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/SNP_kinship/PanGWAS.txt"), sep = "\t", header=T) |> filter(af >=0.05, af <0.95)
# Get significant hits
PanGWAS_SNP_kinship_sighits <- PanGWAS_SNP_kinship_data |> filter(lrt.pvalue < 0.05 / nrow(PanGWAS_SNP_kinship_data)) 
# QQ plot 
PanGWAS_SNP_kinship_qqplot <- QQ_plot(PanGWAS_SNP_kinship_data, "PanGWAS")
```

# 1.5. Unitig_kinship
```{r}
## SNP-GWAS
# Load GWAS results
SNPGWAS_Unitig_kinship_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/Unitig_kinship/SNPGWAS.txt"), sep = "\t", header=T)|> 
  filter(af>=0.05, af <0.95)
# Significant hits
SNPGWAS_Unitig_kinship_sighits <- SNPGWAS_Unitig_kinship_data |>  filter(lrt.pvalue < 0.05 /nrow(SNPGWAS_Unitig_kinship_data)) 
# QQ plot 
SNPGWAS_Unitig_kinship_qqplot <- QQ_plot(SNPGWAS_Unitig_kinship_data, "SNP-GWAS") 

## DBGWAS
# Load GWAS results
DBGWAS_Unitig_kinship_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/Unitig_kinship/DBGWAS.txt"), sep = "\t", header=T) |>
  filter(af >=0.05, af <0.95)
# Significant hits
DBGWAS_Unitig_kinship_sighits <- DBGWAS_Unitig_kinship_data |> filter(lrt.pvalue < 0.05 /87899)  
# QQ plot 
DBGWAS_Unitig_kinship_qqplot <- QQ_plot(DBGWAS_Unitig_kinship_data, "DBGWAS")

## PanGWAS
# Load GWAS results
PanGWAS_Unitig_kinship_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/Unitig_kinship/PanGWAS.txt"), sep = "\t", header=T) |> filter(af >=0.05, af <0.95)
# Get significant hits
PanGWAS_Unitig_kinship_sighits <- PanGWAS_Unitig_kinship_data |> filter(lrt.pvalue < 0.05 / nrow(PanGWAS_Unitig_kinship_data)) 
# QQ plot 
PanGWAS_Unitig_kinship_qqplot <- QQ_plot(PanGWAS_Unitig_kinship_data, "PanGWAS")
```

# 1.6. COG_kinship
```{r}
## SNP-GWAS
# Load GWAS results
SNPGWAS_COG_kinship_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/COG_kinship/SNPGWAS.txt"), sep = "\t", header=T)|> 
  filter(af>=0.05, af <0.95)
# Significant hits
SNPGWAS_COG_kinship_sighits <- SNPGWAS_COG_kinship_data |>  filter(lrt.pvalue < 0.05 /nrow(SNPGWAS_COG_kinship_data)) 
# QQ plot 
SNPGWAS_COG_kinship_qqplot <- QQ_plot(SNPGWAS_COG_kinship_data, "SNP-GWAS") 

## DBGWAS
# Load GWAS results
DBGWAS_COG_kinship_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/COG_kinship/DBGWAS.txt"), sep = "\t", header=T) |>
  filter(af >=0.05, af <0.95)
# Significant hits
DBGWAS_COG_kinship_sighits <- DBGWAS_COG_kinship_data |> filter(lrt.pvalue < 0.05 /87899)  
# QQ plot 
DBGWAS_COG_kinship_qqplot <- QQ_plot(DBGWAS_COG_kinship_data, "DBGWAS")

## PanGWAS
# Load GWAS results
PanGWAS_COG_kinship_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/COG_kinship/PanGWAS.txt"), sep = "\t", header=T) |> filter(af >=0.05, af <0.95)
# Get significant hits
PanGWAS_COG_kinship_sighits <- PanGWAS_COG_kinship_data |> filter(lrt.pvalue < 0.05 / nrow(PanGWAS_COG_kinship_data)) 
# QQ plot 
PanGWAS_COG_kinship_qqplot <- QQ_plot(PanGWAS_COG_kinship_data, "PanGWAS")
```

# 1.7. RAxML
```{r}
## SNP-GWAS
# Load GWAS results
SNPGWAS_RAxML_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/RAxML/SNPGWAS.txt"), sep = "\t", header=T)|> 
  filter(af>=0.05, af <0.95)
# Significant hits
SNPGWAS_RAxML_sighits <- SNPGWAS_RAxML_data |>  filter(lrt.pvalue < 0.05 /nrow(SNPGWAS_RAxML_data)) 
# QQ plot 
SNPGWAS_RAxML_qqplot <- QQ_plot(SNPGWAS_RAxML_data, "SNP-GWAS") 

## DBGWAS
# Load GWAS results
DBGWAS_RAxML_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/RAxML/DBGWAS.txt"), sep = "\t", header=T) |>
  filter(af >=0.05, af <0.95)
# Significant hits
DBGWAS_RAxML_sighits <- DBGWAS_RAxML_data |> filter(lrt.pvalue < 0.05 /87899)  
# QQ plot 
DBGWAS_RAxML_qqplot <- QQ_plot(DBGWAS_RAxML_data, "DBGWAS")

## PanGWAS
# Load GWAS results
PanGWAS_RAxML_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/RAxML/PanGWAS.txt"), sep = "\t", header=T) |> filter(af >=0.05, af <0.95)
# Get significant hits
PanGWAS_RAxML_sighits <- PanGWAS_RAxML_data |> filter(lrt.pvalue < 0.05 / nrow(PanGWAS_RAxML_data)) 
# QQ plot 
PanGWAS_RAxML_qqplot <- QQ_plot(PanGWAS_RAxML_data, "PanGWAS")
```


# 1.8. IQtree
```{r}
## SNP-GWAS
# Load GWAS results
SNPGWAS_IQtree_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/IQtree/SNPGWAS.txt"), sep = "\t", header=T)|> 
  filter(af>=0.05, af <0.95)
# Significant hits
SNPGWAS_IQtree_sighits <- SNPGWAS_IQtree_data |>  filter(lrt.pvalue < 0.05 /nrow(SNPGWAS_IQtree_data)) 
# QQ plot 
SNPGWAS_IQtree_qqplot <- QQ_plot(SNPGWAS_IQtree_data, "SNP-GWAS") 

## DBGWAS
# Load GWAS results
DBGWAS_IQtree_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/IQtree/DBGWAS.txt"), sep = "\t", header=T) |>
  filter(af >=0.05, af <0.95)
# Significant hits
DBGWAS_IQtree_sighits <- DBGWAS_IQtree_data |> filter(lrt.pvalue < 0.05 /87899)  
# QQ plot 
DBGWAS_IQtree_qqplot <- QQ_plot(DBGWAS_IQtree_data, "DBGWAS")

## PanGWAS
# Load GWAS results
PanGWAS_IQtree_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/IQtree/PanGWAS.txt"), sep = "\t", header=T) |> filter(af >=0.05, af <0.95)
# Get significant hits
PanGWAS_IQtree_sighits <- PanGWAS_IQtree_data |> filter(lrt.pvalue < 0.05 / nrow(PanGWAS_IQtree_data)) 
# QQ plot 
PanGWAS_IQtree_qqplot <- QQ_plot(PanGWAS_IQtree_data, "PanGWAS")
```

# 1.9. Fastree SP
```{r}
## SNP-GWAS
# Load GWAS results
SNPGWAS_Fasttree_SP_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/Fasttree_SP/SNPGWAS.txt"), sep = "\t", header=T)|> 
  filter(af>=0.05, af <0.95)
# Significant hits
SNPGWAS_Fasttree_SP_sighits <- SNPGWAS_Fasttree_SP_data |>  filter(lrt.pvalue < 0.05 /nrow(SNPGWAS_Fasttree_SP_data)) 
# QQ plot 
SNPGWAS_Fasttree_SP_qqplot <- QQ_plot(SNPGWAS_Fasttree_SP_data, "SNP-GWAS") 

## DBGWAS
# Load GWAS results
DBGWAS_Fasttree_SP_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/Fasttree_SP/DBGWAS.txt"), sep = "\t", header=T) |>
  filter(af >=0.05, af <0.95)
# Significant hits
DBGWAS_Fasttree_SP_sighits <- DBGWAS_Fasttree_SP_data |> filter(lrt.pvalue < 0.05 /87899)  
# QQ plot 
DBGWAS_Fasttree_SP_qqplot <- QQ_plot(DBGWAS_Fasttree_SP_data, "DBGWAS")

## PanGWAS
# Load GWAS results
PanGWAS_Fasttree_SP_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/Fasttree_SP/PanGWAS.txt"), sep = "\t", header=T) |> filter(af >=0.05, af <0.95)
# Get significant hits
PanGWAS_Fasttree_SP_sighits <- PanGWAS_Fasttree_SP_data |> filter(lrt.pvalue < 0.05 / nrow(PanGWAS_Fasttree_SP_data)) 
# QQ plot 
PanGWAS_Fasttree_SP_qqplot <- QQ_plot(PanGWAS_Fasttree_SP_data, "PanGWAS")
```
# 1.10. Fastree DP FEM
```{r}
## SNP-GWAS
# Load GWAS results
SNPGWAS_Fasttree_DP_FEM_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/Fasttree_DP_FEM/SNPGWAS.txt"), sep = "\t", header=T)|> 
  filter(af>=0.05, af <0.95)
# Significant hits
SNPGWAS_Fasttree_DP_FEM_sighits <- SNPGWAS_Fasttree_DP_FEM_data |>  filter(lrt.pvalue < 0.05 /nrow(SNPGWAS_Fasttree_DP_FEM_data)) 
# QQ plot 
SNPGWAS_Fasttree_DP_FEM_qqplot <- QQ_plot(SNPGWAS_Fasttree_DP_FEM_data, "SNP-GWAS") 

## DBGWAS
# Load GWAS results
DBGWAS_Fasttree_DP_FEM_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/Fasttree_DP_FEM/DBGWAS.txt"), sep = "\t", header=T) |>
  filter(af >=0.05, af <0.95)
# Significant hits
DBGWAS_Fasttree_DP_FEM_sighits <- DBGWAS_Fasttree_DP_FEM_data |> filter(lrt.pvalue < 0.05 /87899)  
# QQ plot 
DBGWAS_Fasttree_DP_FEM_qqplot <- QQ_plot(DBGWAS_Fasttree_DP_FEM_data, "DBGWAS")

## PanGWAS
# Load GWAS results
PanGWAS_Fasttree_DP_FEM_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/Fasttree_DP_FEM/PanGWAS.txt"), sep = "\t", header=T) |> filter(af >=0.05, af <0.95)
# Get significant hits
PanGWAS_Fasttree_DP_FEM_sighits <- PanGWAS_Fasttree_DP_FEM_data |> filter(lrt.pvalue < 0.05 / nrow(PanGWAS_Fasttree_DP_FEM_data)) 
# QQ plot 
PanGWAS_Fasttree_DP_FEM_qqplot <- QQ_plot(PanGWAS_Fasttree_DP_FEM_data, "PanGWAS")
```
# 1.11. Fastree DP LMM
```{r}
## SNP-GWAS
# Load GWAS results
SNPGWAS_Fasttree_DP_LMM_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/Fasttree_DP_LMM/SNPGWAS.txt"), sep = "\t", header=T)|> 
  filter(af>=0.05, af <0.95)
# Significant hits
SNPGWAS_Fasttree_DP_LMM_sighits <- SNPGWAS_Fasttree_DP_LMM_data |>  filter(lrt.pvalue < 0.05 /nrow(SNPGWAS_Fasttree_DP_LMM_data)) |>
  mutate(
    beta_lower = (beta - 1.96 * beta.std.err),
    beta_upper = (beta + 1.96 * beta.std.err),
    beta_CI = sprintf("%.2f (%.2f; %.2f)", beta, beta_lower, beta_upper),
    P_value = formatC(lrt.pvalue, format = "e", digits = 2))
# QQ plot 
SNPGWAS_Fasttree_DP_LMM_qqplot <- QQ_plot(SNPGWAS_Fasttree_DP_LMM_data, "SNP-GWAS") 

## DBGWAS
# Load GWAS results
DBGWAS_Fasttree_DP_LMM_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/Fasttree_DP_LMM/DBGWAS.txt"), sep = "\t", header=T) |>
  filter(af >=0.05, af <0.95)
# Significant hits
DBGWAS_Fasttree_DP_LMM_sighits <- DBGWAS_Fasttree_DP_LMM_data |> 
  filter(lrt.pvalue < 0.05 /87899)  |>
  mutate(
    beta_lower = (beta - 1.96 * beta.std.err),
    beta_upper = (beta + 1.96 * beta.std.err),
    beta_CI = sprintf("%.2f (%.2f; %.2f)", beta, beta_lower, beta_upper),
    P_value = formatC(lrt.pvalue, format = "e", digits = 2))
# QQ plot 
DBGWAS_Fasttree_DP_LMM_qqplot <- QQ_plot(DBGWAS_Fasttree_DP_LMM_data, "DBGWAS")

## PanGWAS
# Load GWAS results
PanGWAS_Fasttree_DP_LMM_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/Fasttree_DP_LMM/PanGWAS.txt"), sep = "\t", header=T) |> filter(af >=0.05, af <0.95)
# Get significant hits
PanGWAS_Fasttree_DP_LMM_sighits <- PanGWAS_Fasttree_DP_LMM_data |> 
  filter(lrt.pvalue < 0.05 / nrow(PanGWAS_Fasttree_DP_LMM_data)) |>
  mutate(
    beta_lower = (beta - 1.96 * beta.std.err),
    beta_upper = (beta + 1.96 * beta.std.err),
    beta_CI = sprintf("%.2f (%.2f; %.2f)", beta, beta_lower, beta_upper),
    P_value = formatC(lrt.pvalue, format = "e", digits = 2)
  )
# QQ plot 
PanGWAS_Fasttree_DP_LMM_qqplot <- QQ_plot(PanGWAS_Fasttree_DP_LMM_data, "PanGWAS")
```
# 1.12. Fastree DP LMM - no recombination
```{r}
## SNP-GWAS
# Load GWAS results
SNPGWAS_Fasttree_DP_noRecombination_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/Fasttree_DP_noRecombination/SNPGWAS.txt"), sep = "\t", header=T)|> 
  filter(af>=0.05, af <0.95)
# Significant hits
SNPGWAS_Fasttree_DP_noRecombination_sighits <- SNPGWAS_Fasttree_DP_noRecombination_data |>
  filter(lrt.pvalue < 0.05 /nrow(SNPGWAS_Fasttree_DP_noRecombination_data))  |>
  mutate(
    beta_lower = (beta - 1.96 * beta.std.err),
    beta_upper = (beta + 1.96 * beta.std.err),
    beta_CI = sprintf("%.2f (%.2f; %.2f)", beta, beta_lower, beta_upper),
    P_value = formatC(lrt.pvalue, format = "e", digits = 2)
  )
# QQ plot 
SNPGWAS_Fasttree_DP_noRecombination_qqplot <- QQ_plot(SNPGWAS_Fasttree_DP_noRecombination_data, "SNP-GWAS") 

## DBGWAS
# Load GWAS results
DBGWAS_Fasttree_DP_noRecombination_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/Fasttree_DP_noRecombination/DBGWAS.txt"), sep = "\t", header=T) |>
  filter(af >=0.05, af <0.95)
# Significant hits
DBGWAS_Fasttree_DP_noRecombination_sighits <- DBGWAS_Fasttree_DP_noRecombination_data |>
  filter(lrt.pvalue < 0.05 /87899)  |>
  mutate(
    beta_lower = (beta - 1.96 * beta.std.err),
    beta_upper = (beta + 1.96 * beta.std.err),
    beta_CI = sprintf("%.2f (%.2f; %.2f)", beta, beta_lower, beta_upper),
    P_value = formatC(lrt.pvalue, format = "e", digits = 2)
  )
# QQ plot 
DBGWAS_Fasttree_DP_noRecombination_qqplot <- QQ_plot(DBGWAS_Fasttree_DP_noRecombination_data, "DBGWAS")

## PanGWAS
# Load GWAS results
PanGWAS_Fasttree_DP_noRecombination_data <- read.table(paste0(path,"pyseer/Population_Structure_Methods/Fasttree_DP_noRecombination/PanGWAS.txt"), sep = "\t", header=T) |> filter(af >=0.05, af <0.95)
# Get significant hits
PanGWAS_Fasttree_DP_noRecombination_sighits <- PanGWAS_Fasttree_DP_noRecombination_data |>
  filter(lrt.pvalue < 0.05 / nrow(PanGWAS_Fasttree_DP_noRecombination_data))  |>
  mutate(
    beta_lower = (beta - 1.96 * beta.std.err),
    beta_upper = (beta + 1.96 * beta.std.err),
    beta_CI = sprintf("%.2f (%.2f; %.2f)", beta, beta_lower, beta_upper),
    P_value = formatC(lrt.pvalue, format = "e", digits = 2)
  )
# QQ plot 
PanGWAS_Fasttree_DP_noRecombination_qqplot <- QQ_plot(PanGWAS_Fasttree_DP_noRecombination_data, "PanGWAS")
```

# 2. SENSITIVTY ANALYSIS : PANGENOME TOOLS
## 2.1.Pan-GWAS : Roary
```{r}
# Load GWAS results
PanGWAS_ROARY <- read.table(paste0(path,"pyseer/Pangenome_Methods/PanGWAS_ROARY.txt"), sep = "\t", header=T) |> 
  filter(af >=0.05, af <0.95)

# Get significant hits
PanGWAS_ROARY_sig <- PanGWAS_ROARY |>
  filter(lrt.pvalue < 0.05 / nrow(PanGWAS_ROARY))  |>
  mutate(
    beta_lower = (beta - 1.96 * beta.std.err),
    beta_upper = (beta + 1.96 * beta.std.err),
    beta_CI = sprintf("%.2f (%.2f; %.2f)", beta, beta_lower, beta_upper),
    P_value = formatC(lrt.pvalue, format = "e", digits = 2)
  ) |>
  select(variant, af, beta_CI, P_value)
PanGWAS_ROARY_sig


# QQ plot 
PanGWAS_ROARY_qqplot <- QQ_plot(PanGWAS_ROARY, "PanGWAS")
```
## 2.2.Pan-GWAS : Panaroo+CLARC
```{r}
# Load GWAS results
PanGWAS_PANAROO_CLARC <- read.table(paste0(path,"pyseer/Pangenome_Methods/PanGWAS_PANAROO_CLARC.txt"), sep = "\t", header=T) |> 
  filter(af >=0.05, af <0.95)

# Get significant hits
PanGWAS_PANAROO_CLARC_sig <- PanGWAS_PANAROO_CLARC |>
  filter(lrt.pvalue < 0.05 / nrow(PanGWAS_PANAROO_CLARC))  |>
  mutate(
    beta_lower = (beta - 1.96 * beta.std.err),
    beta_upper = (beta + 1.96 * beta.std.err),
    beta_CI = sprintf("%.2f (%.2f; %.2f)", beta, beta_lower, beta_upper),
    P_value = formatC(lrt.pvalue, format = "e", digits = 2)
  ) |>
  select(variant, af, beta_CI, P_value)
PanGWAS_PANAROO_CLARC_sig

# QQ plot 
PanGWAS_PANAROO_CLARC_qqplot <- QQ_plot(PanGWAS_PANAROO_CLARC, "PanGWAS")
```
## 2.3.Pan-GWAS: Roary+CLARC
```{r}
# Load GWAS results
PanGWAS_ROARY_CLARC <- read.table(paste0(path,"pyseer/Pangenome_Methods/PanGWAS_ROARY_CLARC.txt"), sep = "\t", header=T) |> 
  filter(af >=0.05, af <0.95)

# Get significant hits
PanGWAS_ROARY_CLARC_sig <- PanGWAS_ROARY_CLARC |>
  filter(lrt.pvalue < 0.05 / nrow(PanGWAS_ROARY_CLARC))  |>
  mutate(
    beta_lower = (beta - 1.96 * beta.std.err),
    beta_upper = (beta + 1.96 * beta.std.err),
    beta_CI = sprintf("%.2f (%.2f; %.2f)", beta, beta_lower, beta_upper),
    P_value = formatC(lrt.pvalue, format = "e", digits = 2)
  ) |>
  select(variant, af, beta_CI, P_value)
PanGWAS_ROARY_CLARC_sig


# QQ plot 
PanGWAS_ROARY_CLARC_qqplot <- QQ_plot(PanGWAS_ROARY_CLARC, "PanGWAS")
```
# TABLE - PANGENOME TOOLS PYSEER RESULTS
```{r}
# sig level
col_0 <- c("Panaroo",  "Panaroo+CLARC", "Roary", "Roary+CLARC")
col_1 <- c(nrow(PanGWAS_Fasttree_DP_LMM_data), nrow(PanGWAS_PANAROO_CLARC), nrow(PanGWAS_ROARY), nrow(PanGWAS_ROARY_CLARC))
col_2 <- 0.05/col_1
col_3 <- c(nrow(PanGWAS_Fasttree_DP_LMM_sighits), nrow(PanGWAS_PANAROO_CLARC_sig), nrow(PanGWAS_ROARY_sig), nrow(PanGWAS_ROARY_CLARC_sig))

# Col 4
col_4 <- c(PanGWAS_Fasttree_DP_LMM_sighits |>  mutate(col4 = paste0(variant, ": af=", af, ", beta (95% CI): ", beta_CI, ", p-value=",P_value)) |> pull(col4)|> paste(collapse = "; "), 
           "",
           PanGWAS_ROARY_sig |>  mutate(col4 = paste0(variant, ": af=", af, ", beta (95% CI): ", beta_CI, ", p-value=",P_value))|> pull(col4) |> paste(collapse = "; "), 
           "")


data.frame(Pangenome_tool = col_0, 
           N_Accessory_genes = col_1,
           Sig_threshold = formatC(col_2, format = "e", digits = 2),
           N_sig = col_3, 
           Sig_hits= col_4)
```



## TABLE Lambda GC 
```{r}
### Lambda GC
# Function
lambdaGC_fun <- function(df){
  # use the LRT p-values from pyseer
p <- df$lrt.pvalue

# drop NAs and p = 0 (if any)
p <- p[!is.na(p) & p > 0]

# convert p-values to chi-square statistics with 1 df
chisq_obs <- qchisq(1 - p, df = 1)

# genomic inflation factor lambda_GC
lambda_gc <- median(chisq_obs) / qchisq(0.5, df = 1)

round(lambda_gc,2)
}


### Arrange df
lambdaGC_df <- data.frame(Method= c("CC", "SC", "Mash distances","SNP kinship", "Unitig kinship", "COG kinship",  "RAxML", "IQtree", "Fasttree SP", "Fasttree DP", "Fasttree DP", "Fasttree DP - no Recombination"), 
           SNPGWAS_lambdaGC= c(lambdaGC_fun(SNPGWAS_CC_data),
                            lambdaGC_fun(SNPGWAS_SC_data) , 
                            lambdaGC_fun(SNPGWAS_mash_distances_data), 
                            lambdaGC_fun(SNPGWAS_SNP_kinship_data), 
                            lambdaGC_fun(SNPGWAS_Unitig_kinship_data), 
                            lambdaGC_fun(SNPGWAS_COG_kinship_data),
                            lambdaGC_fun(SNPGWAS_RAxML_data), 
                            lambdaGC_fun(SNPGWAS_IQtree_data),
                            lambdaGC_fun(SNPGWAS_Fasttree_SP_data) , 
                            lambdaGC_fun(SNPGWAS_Fasttree_DP_FEM_data), 
                            lambdaGC_fun(SNPGWAS_Fasttree_DP_LMM_data),
                            lambdaGC_fun(SNPGWAS_Fasttree_DP_noRecombination_data)), 
           
           PanGWAS_lambdaGC= c(lambdaGC_fun(PanGWAS_CC_data),
                           lambdaGC_fun(PanGWAS_SC_data) , 
                           lambdaGC_fun(PanGWAS_mash_distances_data),
                           lambdaGC_fun(PanGWAS_SNP_kinship_data) , 
                           lambdaGC_fun(PanGWAS_Unitig_kinship_data) , 
                           lambdaGC_fun(PanGWAS_COG_kinship_data), 
                           lambdaGC_fun(PanGWAS_RAxML_data), 
                           lambdaGC_fun(PanGWAS_IQtree_data) , 
                           lambdaGC_fun(PanGWAS_Fasttree_SP_data) ,  
                           lambdaGC_fun(PanGWAS_Fasttree_DP_FEM_data) ,
                           lambdaGC_fun(PanGWAS_Fasttree_DP_LMM_data) , 
                           lambdaGC_fun(PanGWAS_Fasttree_DP_noRecombination_data)), 
           
           DBGWAS_lambdaGC= c(lambdaGC_fun(DBGWAS_CC_data), 
                            lambdaGC_fun(DBGWAS_SC_data) , 
                            lambdaGC_fun(DBGWAS_mash_distances_data) ,
                            lambdaGC_fun(DBGWAS_SNP_kinship_data) , 
                            lambdaGC_fun(DBGWAS_Unitig_kinship_data), 
                            lambdaGC_fun(DBGWAS_COG_kinship_data) , 
                            lambdaGC_fun(DBGWAS_RAxML_data), 
                            lambdaGC_fun(DBGWAS_IQtree_data),
                            lambdaGC_fun(DBGWAS_Fasttree_SP_data) , 
                            lambdaGC_fun(DBGWAS_Fasttree_DP_FEM_data),
                            lambdaGC_fun(DBGWAS_Fasttree_DP_LMM_data),
                            lambdaGC_fun(DBGWAS_Fasttree_DP_noRecombination_data) ))

lambdaGC_df
```

# FIGURE QQ-PLOTs Panel (Figure 2)
```{r}
library(patchwork)

all_plots <- list(SNPGWAS_CC_qqplot + annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[1,2])),  hjust = 1.1, vjust = -0.5,  size = 5), 
                  SNPGWAS_SC_qqplot+ annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[2,2])),  hjust = 1.1, vjust = -0.5,  size = 5), 
                  SNPGWAS_mash_distances_qqplot+ annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[3,2])),  hjust = 1.1, vjust = -0.5,  size = 5), 
                  SNPGWAS_SNP_kinship_qqplot+ annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[4,2])),  hjust = 1.1, vjust = -0.5,  size = 5), 
                  SNPGWAS_Unitig_kinship_qqplot+ annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[5,2])),  hjust = 1.1, vjust = -0.5,  size = 5), 
                  SNPGWAS_COG_kinship_qqplot+ annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[6,2])),  hjust = 1.1, vjust = -0.5,  size = 5), 
                  SNPGWAS_RAxML_qqplot+ annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[7,2])),  hjust = 1.1, vjust = -0.5,  size = 5),
                  SNPGWAS_IQtree_qqplot+ annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[8,2])),  hjust = 1.1, vjust = -0.5,  size = 5), 
                  SNPGWAS_Fasttree_SP_qqplot+ annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[9,2])),  hjust = 1.1, vjust = -0.5,  size = 5), 
                  SNPGWAS_Fasttree_DP_FEM_qqplot+ annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[10,2])),  hjust = 1.1, vjust = -0.5,  size = 5), 
                  SNPGWAS_Fasttree_DP_LMM_qqplot+ annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[11,2])),  hjust = 1.1, vjust = -0.5,  size = 5), 
                  SNPGWAS_Fasttree_DP_noRecombination_qqplot+ annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[12,2])),  hjust = 1.1, vjust = -0.5,  size = 5),
                  
                  DBGWAS_CC_qqplot + annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[1,3])),  hjust = 1.1, vjust = -0.5,  size = 5), 
                  DBGWAS_SC_qqplot+ annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[2,3])),  hjust = 1.1, vjust = -0.5,  size = 5), 
                  DBGWAS_mash_distances_qqplot+ annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[3,3])),  hjust = 1.1, vjust = -0.5,  size = 5), 
                  DBGWAS_SNP_kinship_qqplot+ annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[4,3])),  hjust = 1.1, vjust = -0.5,  size = 5), 
                  DBGWAS_Unitig_kinship_qqplot+ annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[5,3])),  hjust = 1.1, vjust = -0.5,  size = 5), 
                  DBGWAS_COG_kinship_qqplot+ annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[6,3])),  hjust = 1.1, vjust = -0.5,  size = 5), 
                  DBGWAS_RAxML_qqplot+ annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[7,3])),  hjust = 1.1, vjust = -0.5,  size = 5),
                  DBGWAS_IQtree_qqplot+ annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[8,3])),  hjust = 1.1, vjust = -0.5,  size = 5), 
                  DBGWAS_Fasttree_SP_qqplot+ annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[9,3])),  hjust = 1.1, vjust = -0.5,  size = 5), 
                  DBGWAS_Fasttree_DP_FEM_qqplot+ annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[10,3])),  hjust = 1.1, vjust = -0.5,  size = 5), 
                  DBGWAS_Fasttree_DP_LMM_qqplot+ annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[11,3])),  hjust = 1.1, vjust = -0.5,  size = 5), 
                  DBGWAS_Fasttree_DP_noRecombination_qqplot+ annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[12,3])),  hjust = 1.1, vjust = -0.5,  size = 5),
                  
                  
                  PanGWAS_CC_qqplot + annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[1,4])),  hjust = 1.1, vjust = -0.5,  size = 5), 
                  PanGWAS_SC_qqplot+ annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[2,4])),  hjust = 1.1, vjust = -0.5,  size = 5), 
                  PanGWAS_mash_distances_qqplot+ annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[3,4])),  hjust = 1.1, vjust = -0.5,  size = 5), 
                  PanGWAS_SNP_kinship_qqplot+ annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[4,4])),  hjust = 1.1, vjust = -0.5,  size = 5), 
                  PanGWAS_Unitig_kinship_qqplot+ annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[5,4])),  hjust = 1.1, vjust = -0.5,  size = 5), 
                  PanGWAS_COG_kinship_qqplot+ annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[6,4])),  hjust = 1.1, vjust = -0.5,  size = 5), 
                  PanGWAS_RAxML_qqplot+ annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[7,4])),  hjust = 1.1, vjust = -0.5,  size = 5),
                  PanGWAS_IQtree_qqplot+ annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[8,4])),  hjust = 1.1, vjust = -0.5,  size = 5), 
                  PanGWAS_Fasttree_SP_qqplot+ annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[9,4])),  hjust = 1.1, vjust = -0.5,  size = 5), 
                  PanGWAS_Fasttree_DP_FEM_qqplot+ annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[10,4])),  hjust = 1.1, vjust = -0.5,  size = 5), 
                  PanGWAS_Fasttree_DP_LMM_qqplot+ annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[11,4])),  hjust = 1.1, vjust = -0.5,  size = 5), 
                  PanGWAS_Fasttree_DP_noRecombination_qqplot+ annotate("text", x = Inf, y = -Inf, label = substitute(lambda*"GC = "*x, list(x = lambdaGC_df[12,4])),  hjust = 1.1, vjust = -0.5,  size = 5)
                  )

combined_panel <- wrap_plots(all_plots, ncol = 12, nrow = 3)

# Define your 12 titles
titles <- c("CC", "SC", "Mash\nDistances", 
            "SNP\nKinship", "Unitig\nKinship", "COG\nKinship", 
            "RAxML", "IQtree", 
            "Fasttree SP", "Fasttree DP\n(FEM)", "Fasttree DP\n(LMM)", 
            "Fasttree DP\n(Recombination-free)")

# Create title plots
title_plots <- lapply(titles, function(title) {
  ggplot() + 
    annotate("text", x = 0.5, y = 0.5, label = title, 
             size = 5, hjust = 0.5, vjust = 0.5, fontface = "bold") +
    theme_void()
})

# Combine titles into one row
title_row <- wrap_plots(title_plots, ncol = 12)

# Create left side labels
row_labels <- c("SNP-GWAS", "DBGWAS", "PanGWAS")
left_plots <- lapply(row_labels, function(label) {
  ggplot() + 
    annotate("text", x = 0.5, y = 0.5, label = label, 
             size = 5, hjust = 0.5, vjust = 0.5, fontface = "bold",
             angle = 90) +
    theme_void()
})

# Left labels with placeholder for title area
left_title_space <- ggplot() + theme_void()
left_column <- left_title_space / wrap_plots(left_plots, ncol = 1, nrow = 3) + 
  plot_layout(heights = c(1, 3))

# Main panel
main_column <- title_row / wrap_plots(all_plots, ncol = 12, nrow = 3) + 
  plot_layout(heights = c(0.25, 3))

# Combine horizontally
combined_panel <- wrap_plots(list(left_column, main_column), ncol = 2, nrow = 1, widths = c(0.25, 12))


ggsave(paste0(outpath, "QQplot_panel.png"), combined_panel, width = 30, height = 10, dpi = 300)
ggsave(paste0(outpath, "QQplot_panel.tiff"), combined_panel, width = 30, height = 10, dpi = 300)

```

## TABLE Number of significant hits
```{r}
### Arrange df
N_sig_hits_df <- data.frame(Method= c("CC", "SC", "Mash distances", "SNP kinship", "Unitig kinship", "COG kinship", "RAxML", "IQtree", "Fasttree SP", "Fasttree DP", "Fasttree DP", "Fasttree DP - no Recombination"), 
           SNPGWAS_nhits= c(nrow(SNPGWAS_CC_sighits),
                            nrow(SNPGWAS_SC_sighits) , 
                            nrow(SNPGWAS_mash_distances_sighits),
                            nrow(SNPGWAS_SNP_kinship_sighits), 
                            nrow(SNPGWAS_Unitig_kinship_sighits), 
                            nrow(SNPGWAS_COG_kinship_sighits), 
                            nrow(SNPGWAS_RAxML_sighits), 
                            nrow(SNPGWAS_IQtree_sighits),
                            nrow(SNPGWAS_Fasttree_SP_sighits) , 
                            nrow(SNPGWAS_Fasttree_DP_FEM_sighits), 
                            nrow(SNPGWAS_Fasttree_DP_LMM_sighits),
                            nrow(SNPGWAS_Fasttree_DP_noRecombination_sighits)), 
           
           DBGWAS_nhits= c(nrow(DBGWAS_CC_sighits), 
                            nrow(DBGWAS_SC_sighits) , 
                            nrow(DBGWAS_mash_distances_sighits)  ,
                            nrow(DBGWAS_SNP_kinship_sighits) , 
                            nrow(DBGWAS_Unitig_kinship_sighits), 
                            nrow(DBGWAS_COG_kinship_sighits), 
                            nrow(DBGWAS_RAxML_sighits), 
                            nrow(DBGWAS_IQtree_sighits) , 
                            nrow(DBGWAS_Fasttree_SP_sighits) , 
                            nrow(DBGWAS_Fasttree_DP_FEM_sighits),
                            nrow(DBGWAS_Fasttree_DP_LMM_sighits),
                            nrow(DBGWAS_Fasttree_DP_noRecombination_sighits) ), 
           
           PanGWAS_nhits= c(nrow(PanGWAS_CC_sighits),
                           nrow(PanGWAS_SC_sighits) , 
                           nrow(PanGWAS_mash_distances_sighits),
                           nrow(PanGWAS_SNP_kinship_sighits) , 
                           nrow(PanGWAS_Unitig_kinship_sighits) , 
                           nrow(PanGWAS_COG_kinship_sighits), 
                           nrow(PanGWAS_RAxML_sighits) , 
                           nrow(PanGWAS_IQtree_sighits) , 
                           nrow(PanGWAS_Fasttree_SP_sighits) ,  
                           nrow(PanGWAS_Fasttree_DP_FEM_sighits) ,
                           nrow(PanGWAS_Fasttree_DP_LMM_sighits) , 
                           nrow(PanGWAS_Fasttree_DP_noRecombination_sighits)))

N_sig_hits_df 

```

# FIGURE - Correlation between lambdaGC and Number of Sig hits
```{r}
library(dplyr)
library(ggplot2)
library(broom)


# Prepare data
lambdaGC_df_long <- lambdaGC_df %>%
  pivot_longer(
    cols = -Method,
    names_to = "GWAS_type",
    values_to = "lambdaGC"
  ) %>%
  mutate(GWAS_type = str_replace(GWAS_type, "_lambdaGC", ""))

N_sig_hits_df_long <- N_sig_hits_df %>%
  pivot_longer(
    cols = -Method,
    names_to = "GWAS_type",
    values_to = "nhits"
  ) %>%
  mutate(GWAS_type = str_replace(GWAS_type, "_nhits", ""))

lambdaGC_Nsighits_df_long <- left_join(lambdaGC_df_long, N_sig_hits_df_long)

# Get correlation
corr_by_GWAS_type <- lambdaGC_Nsighits_df_long %>%
  group_by(GWAS_type) %>%
  reframe(tidy(cor.test(log10(nhits + 1), lambdaGC, method = "pearson"))) %>%
  mutate(r_lab = round(estimate, 3),
    p_lab = signif(p.value, 3),
    label = paste0("r = ", r_lab, "\nP = ", p_lab))
level_order <- c("SNPGWAS", "PanGWAS", "DBGWAS")
corr_by_GWAS_type2 <- corr_by_GWAS_type %>%
  mutate(GWAS_type = factor(GWAS_type, levels = level_order))

# Plot
plot_corr_lambdaGC_Nsighits <- lambdaGC_Nsighits_df_long  %>%
  ggplot(aes(x = nhits + 1, y = lambdaGC, color = GWAS_type)) +
  geom_point(size=5) +
  facet_grid(. ~ GWAS_type) +
  scale_x_log10() +
  geom_hline(yintercept = 1, linetype = "dashed", color = "red") +
  geom_text(
    data = corr_by_GWAS_type2,
    aes(
      x = 15, y = Inf,
      label = label
    ),
    inherit.aes = FALSE,
    hjust = 1.1, vjust = 1.5,
    size = 3.5
  ) +
  labs(x = "Number of significant hits (log10 scale)",
    y = expression(lambda[GC]),
    color = "GWAS type"
  )+
  scale_color_manual(values= c("lightblue", "darkred", "darkgreen")) +
  theme_classic() +
  theme(legend.position = "none")

plot_corr_lambdaGC_Nsighits

ggsave(paste0(outpath, "Correlation_lambdaGC_Nsighits.png"), plot_corr_lambdaGC_Nsighits, width = 10, height = 5, dpi = 300)
ggsave(paste0(outpath, "Correlation_lambdaGC_Nsighits.tiff"), plot_corr_lambdaGC_Nsighits, width = 10, height = 5, dpi = 300)

```

# TABLE - significant results with and without RECOMBINATION
```{r}
# Recombination
SNPGWAS_Fasttree_DP_LMM_sighits_df <- SNPGWAS_Fasttree_DP_LMM_sighits |> 
  select(variant, af, beta_CI, P_value)|> 
  mutate(POS= str_split(variant, "_", simplify = T)[,2], 
         REF= str_split(variant, "_", simplify = T)[,3],
         ALT= str_split(variant, "_", simplify = T)[,4], 
         variant_new = paste0(REF, POS, ALT))
  
DBGWAS_Fasttree_DP_LMM_sighits_df <- DBGWAS_Fasttree_DP_LMM_sighits |> 
  select(variant, af, beta_CI, P_value)|> 
  mutate(variant_new = paste0("unitig_", 1:nrow(DBGWAS_Fasttree_DP_LMM_sighits)))
 
PanGWAS_Fasttree_DP_LMM_sighits_df <-  PanGWAS_Fasttree_DP_LMM_sighits|> 
  rename(variant_new = variant) |> 
  select(variant_new, af, beta_CI, P_value)|> 
  mutate(annotation="")

# NO Recombination
SNPGWAS_Fasttree_DP_noRecombination_sighits_df <- SNPGWAS_Fasttree_DP_noRecombination_sighits |> 
  select(variant, af, beta_CI, P_value)|> 
  mutate(POS= str_split(variant, "_", simplify = T)[,2], 
         REF= str_split(variant, "_", simplify = T)[,3],
         ALT= str_split(variant, "_", simplify = T)[,4], 
         variant_new = paste0(REF, POS, ALT))
  
DBGWAS_Fasttree_DP_noRecombination_sighits_df <- left_join(DBGWAS_Fasttree_DP_noRecombination_sighits |> select(variant, af, beta_CI, P_value), 
                                                           DBGWAS_Fasttree_DP_LMM_sighits_df |> select(variant, variant_new)) |> 
  arrange(parse_number(variant_new))|>
  mutate(variant_new = ifelse(is.na(variant_new), paste0("unitig_", 21:26, "*"), variant_new))|> 
  arrange(parse_number(variant_new))
  
PanGWAS_Fasttree_DP_noRecombination_sighits_df <-  PanGWAS_Fasttree_DP_noRecombination_sighits|> 
  rename(variant_new = variant) |> 
  select(variant_new, af, beta_CI, P_value)|> 
  mutate(annotation="")


# Load annotations
## SNPGWAS
SNPGWAS_data_Fasttree_LMM_ann <- read.csv(paste0(path,"pyseer/Population_Structure_Methods/Fasttree_DP_LMM/SNP_annotations.csv"), header=F)
colnames(SNPGWAS_data_Fasttree_LMM_ann) <- c("POS", "REF", "ALT", "FUNCLASS", "IMPACT", "gene", "nt_change", "aa_change")
SNPGWAS_data_Fasttree_LMM_ann_df <- SNPGWAS_data_Fasttree_LMM_ann|>
  mutate(variant_new = paste0(REF, POS,  ALT), 
         aa_change = str_remove(aa_change, "p."), 
         type= ifelse(IMPACT == "HIGH", "nonsense", 
                      ifelse(IMPACT == "MODERATE", "missense", 
                             ifelse(IMPACT == "LOW", "synonymous", "non-coding"))), 
         annotation= paste0(type, " (", aa_change, "), gene:", gene))|>
  select(variant_new, annotation)
SNPGWAS_data_Fasttree_LMM_ann_df

SNPGWAS_Fasttree_DP_LMM_sighits_df_final <- left_join(SNPGWAS_Fasttree_DP_LMM_sighits_df, SNPGWAS_data_Fasttree_LMM_ann_df)|> 
  select(variant_new, af, beta_CI, P_value, annotation)

SNPGWAS_Fasttree_DP_noRecombination_sighits_df_final <- left_join(SNPGWAS_Fasttree_DP_noRecombination_sighits_df, SNPGWAS_data_Fasttree_LMM_ann_df)|> 
  select(variant_new, af, beta_CI, P_value, annotation)

## DBGWAS- recombination 
DBGWAS_FastTree_DP_LMM_ann <- read.delim(paste0(path,"pyseer/Population_Structure_Methods/Fasttree_DP_LMM/DBGWAS_annotation.txt"), sep = "\t", header=F) 
colnames(DBGWAS_FastTree_DP_LMM_ann) <- colnames(DBGWAS_Fasttree_DP_LMM_data)
DBGWAS_FastTree_annotations <- DBGWAS_FastTree_DP_LMM_ann |>
  filter(af>0.05, af<0.95)|> 
  mutate(core_accessory = ifelse(str_extract(notes, "^[^:]+")=="AP018935.1", "ref", "unaligned contig"),
         coord_range = str_extract(notes, "(?<=:)[0-9]+-[0-9]+"), 
         start= str_split(coord_range, "-", simplify = T)[,1], 
         end= str_split(coord_range, "-", simplify = T)[,2]) |>
  separate(
    notes, 
    into = c("contig_position", "upstream", "gene", "downstream"), 
    sep = ";", 
    remove = FALSE, 
    fill = "right"
  )|>
  mutate(annotation= paste0("Coord:", start, "-", end, " (", core_accessory, "), gene:",gene))|>
  select(variant, annotation)

DBGWAS_Fasttree_DP_LMM_sighits_df_final <- left_join(DBGWAS_Fasttree_DP_LMM_sighits_df, DBGWAS_FastTree_annotations)|> 
  select(variant_new, af, beta_CI, P_value, annotation)


## DBGWAS- no recombination 
DBGWAS_FastTree_DP_LMM_noRecombination_ann <- read.delim(paste0(path,"pyseer/Population_Structure_Methods/Fasttree_DP_noRecombination/DBGWAS_annotation.txt"), sep = "\t", header=F) 
colnames(DBGWAS_FastTree_DP_LMM_noRecombination_ann) <- colnames(DBGWAS_Fasttree_DP_LMM_data)
DBGWAS_FastTree_noRecombination_annotations <- DBGWAS_FastTree_DP_LMM_noRecombination_ann |>
  filter(af>0.05, af<0.95)|> 
  mutate(core_accessory = ifelse(str_extract(notes, "^[^:]+")=="AP018935.1", "ref", "unaligned contig"),
         coord_range = str_extract(notes, "(?<=:)[0-9]+-[0-9]+"), 
         start= str_split(coord_range, "-", simplify = T)[,1], 
         end= str_split(coord_range, "-", simplify = T)[,2]) |>
  separate(
    notes, 
    into = c("contig_position", "upstream", "gene", "downstream"), 
    sep = ";", 
    remove = FALSE, 
    fill = "right"
  )|>
  mutate(annotation= paste0("Coord:", start, "-", end, " (", core_accessory, "), gene:",gene))|>
  select(variant, annotation)

DBGWAS_Fasttree_DP_noRecombination_sighits_df_final <- left_join(DBGWAS_Fasttree_DP_noRecombination_sighits_df, DBGWAS_FastTree_noRecombination_annotations)|> 
  select(variant_new, af, beta_CI, P_value, annotation)



# Make table for DP with RECOMBINATION
Fasttree_DP_LMM_df <- rbind(SNPGWAS_Fasttree_DP_LMM_sighits_df_final, 
      DBGWAS_Fasttree_DP_LMM_sighits_df_final, 
      PanGWAS_Fasttree_DP_LMM_sighits_df)
Fasttree_DP_LMM_df

Fasttree_DP_LMM_noRecombination_df <- rbind(SNPGWAS_Fasttree_DP_noRecombination_sighits_df_final, 
      DBGWAS_Fasttree_DP_noRecombination_sighits_df_final, 
      PanGWAS_Fasttree_DP_noRecombination_sighits_df)
Fasttree_DP_LMM_noRecombination_df
```


# Make supplementary figure for MS2: methods sensitivity analyses
```{r}
library(dplyr)
library(openxlsx)

wb <- createWorkbook()

# CCs_FEM --------------------------------------------------------------
snpgwas_CCs_FEM <- SNPGWAS_CC_data |>
  filter(lrt.pvalue < 0.05 / nrow(SNPGWAS_CC_data)) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err, notes)


addWorksheet(wb, "SNPGWAS CCs (FEM)")
writeData(wb, "SNPGWAS CCs (FEM)", snpgwas_CCs_FEM)

dbgwass_CCs_FEM <- DBGWAS_CC_data |>
  filter(lrt.pvalue < 0.05 / 87899) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err, notes)


addWorksheet(wb, "DBGWAS CCs (FEM)")
writeData(wb, "DBGWAS CCs (FEM)", dbgwass_CCs_FEM)

pangwas_CCs_FEM <- PanGWAS_CC_data |>
  filter(lrt.pvalue < 0.05 / nrow(PanGWAS_CC_data)) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err, notes)

addWorksheet(wb, "PanGWAS CCs (FEM)")
writeData(wb, "PanGWAS CCs (FEM)", pangwas_CCs_FEM)

# SCs_FEM --------------------------------------------------------------
snpgwas_SCs_FEM <- SNPGWAS_SC_data |>
  filter(lrt.pvalue < 0.05 / nrow(SNPGWAS_SC_data)) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err, notes)

addWorksheet(wb, "SNPGWAS SCs (FEM)")
writeData(wb, "SNPGWAS SCs (FEM)", snpgwas_SCs_FEM)

dbgwass_SCs_FEM <- DBGWAS_SC_data |>
  filter(lrt.pvalue < 0.05 / 87899) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err, notes)


addWorksheet(wb, "DBGWAS SCs (FEM)")
writeData(wb, "DBGWAS SCs (FEM)", dbgwass_SCs_FEM)

pangwas_SCs_FEM <- PanGWAS_SC_data |>
  filter(lrt.pvalue < 0.05 / nrow(PanGWAS_SC_data)) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err, notes)

addWorksheet(wb, "PanGWAS SCs (FEM)")
writeData(wb, "PanGWAS SCs (FEM)", pangwas_SCs_FEM)

# Mash_dist_FEM --------------------------------------------------------
snpgwas_mash_FEM <- SNPGWAS_mash_distances_data |>
  filter(lrt.pvalue < 0.05 / nrow(SNPGWAS_mash_distances_data)) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err, notes)


addWorksheet(wb, "SNPGWAS Mash dist (FEM)")
writeData(wb, "SNPGWAS Mash dist (FEM)", snpgwas_mash_FEM)

dbgwass_mash_FEM <- DBGWAS_mash_distances_data |>
  filter(lrt.pvalue < 0.05 / 87899) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err, notes)

addWorksheet(wb, "DBGWAS Mash dist (FEM)")
writeData(wb, "DBGWAS Mash dist (FEM)", dbgwass_mash_FEM)

pangwas_mash_FEM <- PanGWAS_mash_distances_data |>
  filter(lrt.pvalue < 0.05 / nrow(PanGWAS_mash_distances_data)) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err, notes)

addWorksheet(wb, "PanGWAS Mash dist (FEM)")
writeData(wb, "PanGWAS Mash dist (FEM)", pangwas_mash_FEM)

# RAxML_SNP_phylo_LMM --------------------------------------------------
snpgwas_RAxml <- SNPGWAS_RAxML_data |>
  filter(lrt.pvalue < 0.05 / nrow(SNPGWAS_RAxML_data)) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err,variant_h2, notes)

addWorksheet(wb, "SNPGWAS RaxML (LMM)")
writeData(wb, "SNPGWAS RaxML (LMM)", snpgwas_RAxml)

dbgwass_RAxml <- DBGWAS_RAxML_data |>
  filter(lrt.pvalue < 0.05 /87899) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err,variant_h2, notes)

addWorksheet(wb, "DBGWAS RaxML (LMM)")
writeData(wb, "DBGWAS RaxML (LMM)", dbgwass_RAxml)

pangwas_RAxml <- PanGWAS_RAxML_data |>
  filter(lrt.pvalue < 0.05 / nrow(PanGWAS_RAxML_data)) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err,variant_h2, notes)

addWorksheet(wb, "PanGWAS RaxML (LMM)")
writeData(wb, "PanGWAS RaxML (LMM)", pangwas_RAxml)

# IQtree_SNP_phylo_LMM -------------------------------------------------
snpgwas_IQtree <- SNPGWAS_IQtree_data |>
  filter(lrt.pvalue < 0.05 / nrow(SNPGWAS_IQtree_data)) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err,variant_h2, notes)


addWorksheet(wb, "SNPGWAS IQtree (LMM)")
writeData(wb, "SNPGWAS IQtree (LMM)", snpgwas_IQtree)

dbgwass_IQtree <- DBGWAS_IQtree_data |>
  filter(lrt.pvalue < 0.05 /87899) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err,variant_h2, notes)


addWorksheet(wb, "DBGWAS IQtree (LMM)")
writeData(wb, "DBGWAS IQtree (LMM)", dbgwass_IQtree)

pangwas_IQtree <- PanGWAS_IQtree_data |>
  filter(lrt.pvalue < 0.05 / nrow(PanGWAS_IQtree_data)) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err,variant_h2, notes)

addWorksheet(wb, "PanGWAS IQtree (LMM)")
writeData(wb, "PanGWAS IQtree (LMM)", pangwas_IQtree)

# Fasttree_SP_phylo_LMM -----------------------------------------------
snpgwas_Fasttree_SP <- SNPGWAS_Fasttree_SP_data |>
  filter(lrt.pvalue < 0.05 / nrow(SNPGWAS_Fasttree_SP_data)) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err,variant_h2, notes)


addWorksheet(wb, "SNPGWAS FastTree SP (LMM)")
writeData(wb, "SNPGWAS FastTree SP (LMM)", snpgwas_Fasttree_SP)

dbgwass_Fasttree_SP <- DBGWAS_Fasttree_SP_data |>
  filter(lrt.pvalue < 0.05 / 87899) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err,variant_h2, notes)


addWorksheet(wb, "DBGWAS FastTree SP (LMM)")
writeData(wb, "DBGWAS FastTree SP (LMM)", dbgwass_Fasttree_SP)

pangwas_Fasttree_SP <- PanGWAS_Fasttree_SP_data |>
  filter(lrt.pvalue < 0.05 / nrow(PanGWAS_Fasttree_SP_data)) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err,variant_h2, notes)

addWorksheet(wb, "PanGWAS FastTree SP (LMM)")
writeData(wb, "PanGWAS FastTree SP (LMM)", pangwas_Fasttree_SP)

# Fasttree_DP_phylo_LMM -----------------------------------------------
snpgwas_Fasttree_DP <- SNPGWAS_Fasttree_DP_LMM_data |>
  filter(lrt.pvalue < 0.05 / nrow(SNPGWAS_Fasttree_DP_LMM_data)) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err,variant_h2, notes)


addWorksheet(wb, "SNPGWAS FastTree DP (LMM)")
writeData(wb, "SNPGWAS FastTree DP (LMM)", snpgwas_Fasttree_DP)

dbgwass_Fasttree_DP <- DBGWAS_Fasttree_DP_LMM_data |>
  filter(lrt.pvalue < 0.05 /87899) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err,variant_h2, notes)

addWorksheet(wb, "DBGWAS FastTree DP (LMM)")
writeData(wb, "DBGWAS FastTree DP (LMM)", dbgwass_Fasttree_DP)

pangwas_Fasttree_DP <- PanGWAS_Fasttree_DP_LMM_data |>
  filter(lrt.pvalue < 0.05 / nrow(PanGWAS_Fasttree_DP_LMM_data)) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err,variant_h2, notes)

addWorksheet(wb, "PanGWAS FastTree DP (LMM)")
writeData(wb, "PanGWAS FastTree DP (LMM)", pangwas_Fasttree_DP)

# Fasttree_DP_phylo_FEM -----------------------------------------------
snpgwas_Fasttree_DP_FEM <- SNPGWAS_Fasttree_DP_FEM_data |>
  filter(lrt.pvalue < 0.05 / nrow(SNPGWAS_Fasttree_DP_FEM_data)) |>
   select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err, notes)


addWorksheet(wb, "SNPGWAS FastTree DP (FEM)")
writeData(wb, "SNPGWAS FastTree DP (FEM)", snpgwas_Fasttree_DP_FEM)

dbgwass_Fasttree_DP_FEM <- DBGWAS_Fasttree_DP_FEM_data |>
  filter(lrt.pvalue < 0.05 /87899) |>
   select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err, notes)


addWorksheet(wb, "DBGWAS FastTree DP (FEM)")
writeData(wb, "DBGWAS FastTree DP (FEM)", dbgwass_Fasttree_DP_FEM)

pangwas_Fasttree_DP_FEM <- PanGWAS_Fasttree_DP_FEM_data |>
  filter(lrt.pvalue < 0.05 / nrow(PanGWAS_Fasttree_DP_FEM_data)) |>
   select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err, notes)

addWorksheet(wb, "PanGWAS FastTree DP (FEM)")
writeData(wb, "PanGWAS FastTree DP (FEM)", pangwas_Fasttree_DP_FEM)

# SNP_kinship_LMM ----------------------------------------------
snpgwas_SNP_kinship <- SNPGWAS_SNP_kinship_data |>
  filter(lrt.pvalue < 0.05 / nrow(SNPGWAS_SNP_kinship_data)) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err,variant_h2, notes)


addWorksheet(wb, "SNPGWAS SNP kinship (LMM)")
writeData(wb, "SNPGWAS SNP kinship (LMM)", snpgwas_SNP_kinship)

dbgwass_SNP_kinship <- DBGWAS_SNP_kinship_data |>
  filter(lrt.pvalue < 0.05 /87899) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err,variant_h2, notes)

addWorksheet(wb, "DBGWAS SNP kinship (LMM)")
writeData(wb, "DBGWAS SNP kinship (LMM)", dbgwass_SNP_kinship)

pangwas_SNP_kinship <- PanGWAS_SNP_kinship_data |>
  filter(lrt.pvalue < 0.05 / nrow(PanGWAS_SNP_kinship_data)) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err,variant_h2, notes)

addWorksheet(wb, "PanGWAS SNP kinship (LMM)")
writeData(wb, "PanGWAS SNP kinship (LMM)", pangwas_SNP_kinship)

# Unitig_kinship_LMM -------------------------------------------
snpgwas_unitig_kinship <- SNPGWAS_Unitig_kinship_data |>
  filter(lrt.pvalue < 0.05 / nrow(SNPGWAS_Unitig_kinship_data)) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err,variant_h2, notes)

addWorksheet(wb, "SNPGWAS Unitig kinship (LMM)")
writeData(wb, "SNPGWAS Unitig kinship (LMM)", snpgwas_unitig_kinship)

dbgwass_unitig_kinship <- DBGWAS_Unitig_kinship_data |>
  filter(lrt.pvalue < 0.05 /87899) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err,variant_h2, notes)

addWorksheet(wb, "DBGWAS Unitig kinship (LMM)")
writeData(wb, "DBGWAS Unitig kinship (LMM)", dbgwass_unitig_kinship)

pangwas_unitig_kinship <- PanGWAS_Unitig_kinship_data |>
  filter(lrt.pvalue < 0.05 / nrow(PanGWAS_Unitig_kinship_data)) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err,variant_h2, notes)

addWorksheet(wb, "PanGWAS Unitig kinship (LMM)")
writeData(wb, "PanGWAS Unitig kinship (LMM)", pangwas_unitig_kinship)

# COG_kinship_LMM ----------------------------------------------
snpgwas_COG_kinship <- SNPGWAS_COG_kinship_data |>
  filter(lrt.pvalue < 0.05 / nrow(SNPGWAS_COG_kinship_data)) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err,variant_h2, notes)

addWorksheet(wb, "SNPGWAS COG kinship (LMM)")
writeData(wb, "SNPGWAS COG kinship (LMM)", snpgwas_COG_kinship)

dbgwass_COG_kinship <- DBGWAS_COG_kinship_data |>
  filter(lrt.pvalue < 0.05 /87899) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err,variant_h2, notes)

addWorksheet(wb, "DBGWAS COG kinship (LMM)")
writeData(wb, "DBGWAS COG kinship (LMM)", dbgwass_COG_kinship)

pangwas_COG_kinship <- PanGWAS_COG_kinship_data |>
  filter(lrt.pvalue < 0.05 / nrow(PanGWAS_COG_kinship_data)) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err,variant_h2, notes)

addWorksheet(wb, "PanGWAS COG kinship (LMM)")
writeData(wb, "PanGWAS COG kinship (LMM)", pangwas_COG_kinship)

# Recombination_free_Fasttree_DP_phylo_LMM ----------------------------
snpgwas_gubbins <- SNPGWAS_Fasttree_DP_noRecombination_data |>
  filter(lrt.pvalue < 0.05 / nrow(SNPGWAS_Fasttree_DP_noRecombination_data)) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err,variant_h2, notes)

addWorksheet(wb, "SNPGWAS No recomb (LMM)")
writeData(wb, "SNPGWAS No recomb (LMM)", snpgwas_gubbins)

dbgwass_gubbins <- DBGWAS_Fasttree_DP_noRecombination_data |>
  filter(lrt.pvalue < 0.05 /87899) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err,variant_h2, notes)

addWorksheet(wb, "DBGWAS No recomb (LMM)")
writeData(wb, "DBGWAS No recomb (LMM)", dbgwass_gubbins)

pangwas_gubbins <- PanGWAS_Fasttree_DP_noRecombination_data |>
  filter(lrt.pvalue < 0.05 / nrow(PanGWAS_Fasttree_DP_noRecombination_data)) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err,variant_h2, notes)

addWorksheet(wb, "PanGWAS No recomb (LMM)")
writeData(wb, "PanGWAS No recomb (LMM)", pangwas_gubbins)

# Pangenome tools ------------------------------------------------------
# Roary
pangwas_Roary <- PanGWAS_ROARY |>
  filter(lrt.pvalue < 0.05 / nrow(PanGWAS_ROARY)) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err,variant_h2, notes)

addWorksheet(wb, "PanGWAS Roary")
writeData(wb, "PanGWAS Roary", pangwas_Roary)

# Panaroo_CLARC
pangwas_Panaroo_CLARC <- PanGWAS_PANAROO_CLARC |>
  filter(lrt.pvalue < 0.05 / nrow(PanGWAS_PANAROO_CLARC)) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err,variant_h2, notes)

addWorksheet(wb, "PanGWAS Panaroo + CLARC")
writeData(wb, "PanGWAS Panaroo + CLARC", pangwas_Panaroo_CLARC)

# Roary_CLARC
pangwas_Roary_CLARC <- PanGWAS_ROARY_CLARC |>
  filter(lrt.pvalue < 0.05 / nrow(PanGWAS_ROARY_CLARC)) |>
  select(variant, af, filter.pvalue, lrt.pvalue, beta, beta.std.err,variant_h2, notes)

addWorksheet(wb, "PanGWAS Roary + CLARC")
writeData(wb, "PanGWAS Roary + CLARC", pangwas_Roary_CLARC)

# Finally save workbook ------------------------------------------------
saveWorkbook(wb, paste0(outpath, "Supp_Pyseer_significant_outputs"), overwrite = TRUE)
```




